Simulating dynein's powerstroke using Brownian dynamics

### 2017-05-12 thesis/latex/thesis.tex
Thesis: Add work to introduction.

Cianfrocco \textit{et. al} \cite{cianfroccoreview} propose dynein achieves motion via transitioning through the cycle of states shown in Fig. \ref{mech-cycle}. This model will hereon be referred to as the Cianfrocco model. The model claims the primary events in the dynein step are as follows. First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation. ATP hydrolysis to ADP-Pi causes further conformation changes. These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound. The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT. Finally the binding domain rebinds to the MT, causing a final conformation change in the motor which returns it to the beginning of the cycle. The exact order of these events is not precisely known.\\

@@

\section{This work}
\subsection{Necessary and sufficient criteria for motion}
One of the principle goals of researchers studying the dynein motor is establishing how it walks. This can be a difficult task, since direct observation of the motor walking is not currently experimentally possible. The next best thing, then, is building a model which both only does all the things dynein does, and generates motion. Finding such a system would be very useful for forming hypotheses on how dynein walks. Another requirement of the system would be that it be as simple as possible while still explaining all the data. By Occam's Razor, such systems are the most likely to represent real dynein. Thus, finding the necessary and sufficient criteria for dynein's motion is a big goal in the field.\\

A good starting place for finding the minimum set of motile criteria would be preexisting models, like the Cianfrocco model. This model describes a set of states which logically lead to each other and are supported by experiment. It seems intuitive that a motor which obeyed the Cianfrocco cycle would walk. There are multiple ways to verify this. A completely synthetic biological motor could be designed which obeys the cycle, and its motile properties could be studied. This seems incredibly difficult. Much easier would be to create a physical model of such a cycle, simulate it, and see if it obeys similarly to experimentally observed dynein. If the physical model both behaves according to the laws of physics and captures experimental dynein behavior well, this would be good evidence that such a model describes how dynein actually walks.\\

\subsection{Goals}
The goal of this work is to construct a minimum physical model of dynein protein based on the theoretical Cianfrocco model. The model should behave as physics predicts a protein to behave; namely, it should obey Brownian dynamics and conservation of energy. It should also roughly capture the state transitions and conformation changes of the Cianfrocco model. Also the model should be as simple as possible, to assure that each part of the model is necessary. If too many extra features are added, the model will not be minimal and not achieve its purpose.\\

This model will then be converted into a computer program, tested for compliance with the laws of physics, and simulated. Information on the long-term behavior of such a simulation will then be used to evaluate how well the model describe's experimental dynein's behavior.\\

>>>>

Cianfrocco \textit{et. al} \cite{cianfroccoreview} propose dynein achieves motion via transitioning through the cycle of states shown in Fig. \ref{mech-cycle}. The model claims the primary events in the dynein step are as follows. First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation. ATP hydrolysis to ADP-Pi causes further conformation changes. These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound. The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT. Finally the binding domain rebinds to the MT, causing a final conformation change in the motor which returns it to the beginning of the cycle. The exact order of these events is not precisely known.\\

@@

\section{Testing the Cianfrocco model}
One of the main goals of studying dynein is establishing how the motor walks. This can be a difficult task, since direct observation of the protein walking is not currently experimentally possible. Given the experimental challenges, a computational approach seems promising. This approach involves creating a hypothesis on how the motor walks, creating a model which behaves according to this hypothesis, simulating that model, and comparing the simulation's behavior to experimental data. This process can be an efficient way to support hypotheses and guide future research.\\

\subsection{Computational approach}
This project seeks to create a dynamic model of dynein and use it to answer questions. The Cianfrocco model provides a rough framework for the events that go on during the dynein mechanochemical cycle. By augmenting this model with information about dynein's structure and conformation changes occuring during the cycle, a new dynamic model which takes into account both temporal and spatial information can be constructed. This dynamic model can then be simulated under cellular conditions to test how it behaves. Such a model would be ideal, since it would make very precise predictions about how real dynein walks. This simulation would act as an experiment testing the validity of the Cianfrocco model.\\

\subsection{Questions to answer}
The biggest question such a simulation would answer is: does the Cianfrocco model, when applied to a dynein-sized motor with elastic hinges, produce dynein-like stepping? High similarity between simulation behavior and experimental dynein behavior would be a strong positive result supporting the Cianfrocco model. If confirmed to be dynein-like, the simulation could then be used to answer many other questions.\\

Dynein has both angular and linear elasticity between its domains. Angular elasticity, or the flexibility of protein domains relative to each other, may play a large role in dynein's stepping. It is possible that dynein requires its foot domain to diffuse to a high-energy conformation in order to take a step. It is postulated by \textit{Burgess et. al.} \cite{burgess-paper} that changes in angular elasticity may be important for dynein's ability to step at key points in the mechanochemical cycle. Alternatively, it is possible that elasticity is not required, and the angular changes caused by the powerstroke are enough for stepping. By varying the angular elasticity of the model, this simulation could determine the importance of angular elasticity for dynein's step.\\

Another question this simulation could answer is how dynein localizes to its binding site. Microtubules are regular tubulin polymers with MTBD binding sites roughly every 8nm. Dynein can take steps at integer multiples of this length, but the 8nm step is its most favored. Why is this? Does the MTBD drag along the microtubule until it hits a binding site? Or, does it diffuse high above the microtubule, randomly descending until it finds a place to bind. This simulation could answer that question by outputting MTBD positions over time.\\

Why does dynein step differently than kinesin? Kinesin follows a similar mechanochemical cycle to dynein. By modifying the stepping cycle, shrinking inter-domain distance and changing conformation angles of the dynein model, it may be possible to create a rough kinesin dynamic model. This model could then be tested for similarity to experimental kinesin. If kinesin behavior was replicable, then comparisons between the models could be readily made. Does kinesin step like dynein when its size is increased? Does dynein step like kinesin when its conformation angles are made rigid? By varying the spatial representation of the two models, it might be possible to determine why the two motors step differently.\\

Kinesin and dynein generally travel in opposite directions along the microtubule. Why is this? Both proteins bind to the same place on the microtubule \cite{kinesin-dynein-mt-overlap}, so the binding site can't determine directionality. It is possible that not the binding site, but the angle at which the protein binds to the microtubule is what determines its stepping direction. If it turns out that kinesin can be successfully modelled using the same dynamic model, then this question could possibly be answered. If altering the microtubule binding angles of dynein and kinesin reverses their directionality, it may be that binding angle is all it takes to determine motor directionality.\\
